The Hosting Environment of the Advanced Resource Connector middleware

### 2008-08-29 doc/hed/ARCHED_article.tex
Converting HED document into NorduGrid Tech. Doc. according to template. Adding more text about MCCs.

\documentclass[a4paper,10pt]{article}

\usepackage{ifthen}
\newboolean{brieftechdesc}
\newboolean{verbtechdesc}
\setboolean{brieftechdesc}{false}
\setboolean{verbtechdesc}{true}

%opening
\title{The Hosting Environment of the Advanced Resource Connector middleware}

\author{J. J\"{o}nemo, et al}

\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction}

@@

\section{Architecture}

\subsection{Requirements}
% Perhaps this can be an introductory part of the section without any subsection, we'll see
In the design of the HED, several goals and requirements were considered. These were weighed against each other and the factual context.

@@

%\subsection{Functional design}

@@

\subsection{Technical design}

@@

\subsubsection{MCC}

In the HED data channels to the outside world may be set up by chains of small processing units called Message Chain Components (MCCs). These work on units called Messages which represent data going in to or out of the HED. The message consists of the so called Payload which is its main content structured in a way relevant to the protocol of the corresponding MCC, and auxiliary structures such as general Attributes and Security Attributes where information relevant to each protocol is accumulated as the message progresses. Each MCC typically implements one level in the Internet Protocol suite by transforming a message to an input suitable to propagate to the next component and then performs the corresponding transformation of the response on the way back. The components are all dynamically loaded to provide maximum flexibility and extensibility. Each instance of these MCC's can be individually configured.

Each MCC has an entry method which is called with Message being processed. It then processes Message by modifying it or creating new Message. Then MCC calls entry method of next MCC in the chain. For information how mwssages are handled and about memory management policies please see API description of MCC class in [].

@@

\paragraph{Payload}

@@

\paragraph{Attributes}

@@

By convention keys are composed of two parts: name of MCC/Protocol at which Attribute was generated or must be consumed and name of Attribute itself separated by column. For example the Attribute with key HTTP:METHOD holds HTTP protocol method like GET, PUT, HEAD, etc. It is either generated MCC implementing HTTP protocol or is filled by other code and is used by HTTP MCC to generate proper HTTP header.

Each MCC developed inside and outside the HED must be have generated and consumed Attributes described in accompanying documentation.

\paragraph{Security Attributes}

@@

\paragraph{Security Handler}

@@

\paragraph{Client and Server MCCs}

\paragraph{Configuration}

The chain formulation using MCCs is up to the administrator. The chain is an ordered list of MCCs and their interconection can be described in the configuration file. Based on the configuration information the Loader components formulate the actual chain in the memory of the daemon. The MCC may implement some routing algorithm which means one MCC may have connections to multiple other MCCs. Typical scenario is that the HTTP MCC at the server side routes the HTTP messages with POST HTTP operation to a SOAP MCC but the messages with GET operation to for example a simple HTTPD service component.

Most of the MCCs has a client and a server version because the behavior of an MCC should be different depending on whether it is sitting on the server or client side. The typical scenario here is illustrated by the TCP MCC which should listen and wait for incomming messages on a socket on the server side but call \texttt{connect} on the client side.

The developer who writes an MCC is free to choose any 3rd party library and component to implement the functionalty of the MCC but at least currently the MCCs should be written in the same language as the HED was written (C++) and should use the \texttt{MCC} interface class and \texttt{Message} class provided by libarcloader and libarcmessage libraries of ARC1.

\paragraph{TCP MCC}

@@

\paragraph{TLS MCC}
The server TLS MCC enables ...

@@

\paragraph{HTTP MCC}

\paragraph{SOAP MCC}

\paragraph{Plexer}

In general case multiple service living in the HED so the incomming message should route to the proper service. The Plexer MCC does this job. It takes the the ENDPOINT attributes of the message collected by other MCCs compares this attribute to regular expression defined in the configuration file and forward the message to the all matching service. It acts as a dispacher. The Plexer is special in a sence that it is not a plugin but part of the libarcloader library.

\subsubsection{Services}

The services are dynamically loaded on start up just like the MCCs. They can even be regarded as special cases of MCCs in that they constitute the last link in the so called Message Chain and are attached in much the same manner as other MCCs. After being loaded, instanciated and configured by the HED, services are provided with a range of...

The services are typically expected to receive parsed SOAP messages which they service - possibly by invoking or communicating with external processes.

\subsubsection{PDC}

\subsubsection{Alternative implementation languages}

In order to facilitate the developement of services, API bindings for languages other than C++ are provided and some service developement has already been done 
%Not much aboút the actual services, more about what they need to implement and what 
\section{Conclusion}

>>>>

\documentclass{book}
%\documentclass{article}                            %for shorter notes
\usepackage{graphicx}                              %for PNG images (pdflatex)
%\usepackage{graphics}                              %for EPS images (latex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
\usepackage{ifthen}
\newboolean{brieftechdesc}
\newboolean{verbtechdesc}
\setboolean{brieftechdesc}{false}
\setboolean{verbtechdesc}{true}
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {J. J\"{o}nemo, et al},
  pdftitle = {The Hosting Environment of the Advanced Resource Connector middleware},
  pdfsubject = {Paper subject},
  pdfkeywords = {HED,ARC},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\begin{document}
\def\today{\number\day/\number\month/\number\year}

\begin{titlepage}

\begin{tabular}{rl}
\resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
&\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
\end{tabular}

\hrulefill

%-------- Change this to NORDUGRID-XXXXXXX-NN

{\raggedleft NORDUGRID-XXXXXXX-NN\par}

{\raggedleft \today\par}

\vspace*{2cm}

%%%%---- The title ----
{\centering \textsc{\Large The Hosting Environment of the Advanced Resource Connector middleware}\Large \par}
\vspace*{0.5cm}

%%%%---- A subtitle, if necessary ----
%{\centering \textit{\large Paper subtitle}\large \par}

\vspace*{1.5cm}
%%%%---- A list of authors ----
    {\centering \large J. J\"{o}nemo\footnote{authors@address}, et al \large \par}

%%%%---- An abstract - if style is article ----
%\begin{abstract}
%The abstract
%\end{abstract}
\end{titlepage}

\tableofcontents                          %Comment if use article style
\newpage
%\chapter{Preface}
%\section{Introduction}                    %Use Sections for articles
%\label{sec:intro}

\chapter{Introduction}

@@

\chapter{Architecture}

\section{Requirements}
% Perhaps this can be an introductory part of the section without any subsection, we'll see
In the design of the HED, several goals and requirements were considered. These were weighed against each other and the factual context.

@@

%\section{Functional design}

@@

\section{Technical design}

@@

\subsection{MCC}

In the HED data channels to the outside world may be set up by chains of small processing units called Message Chain Components (MCCs). The chain is an ordered list of MCCs and their interconection can be described in the configuration file. The MCCs work on units called Messages which represent data going in to or out of the HED. The message consists of the so called Payload which is its main content structured in a way relevant to the protocol of the corresponding MCC, and auxiliary structures such as general Attributes and Security Attributes where information relevant to each protocol is accumulated as the message progresses. Each MCC typically implements one level in the Internet Protocol suite by transforming a message to an input suitable to propagate to the next component and then performs the corresponding transformation of the response on the way back. The components are all dynamically loaded to provide maximum flexibility and extensibility. Each instance of these MCC's can be individually configured.

Each MCC has an entry method \texttt{process} which is called with Message being processed. It then processes Message by modifying it or creating new Message. Then MCC calls entry method of next MCC in the chain. For information how messages are handled and about memory management policies please see API description of MCC class in [].

The developer who writes an MCC is free to choose any 3rd party library and component to implement the functionalty of the MCC but at least currently the MCCs should be written in the same language as the HED was written (C++) and should use the \texttt{MCC} interface class and \texttt{Message} class provided by \texttt{arcloader} and \texttt{arcmessage} libraries of ARC1.

The MCC may implement some routing algorithm which means one MCC may have connections to multiple other MCCs. Typical scenario is that the HTTP MCC at the server side routes the HTTP messages with POST HTTP operation to a SOAP MCC but the messages with GET operation to for example a simple HTTPD service component.

@@

\subsubsection{Server and Client Side MCCs}

Most of the MCCs has a client and a server version because the behavior of an MCC should be different depending on whether it is sitting on the server or client side. The typical scenario here is illustrated by the TCP MCC which should listen and wait for incomming messages on a socket on the server side but call \texttt{connect()} on the client side.

Server and Client side MCCs are separate elements although definitely sharing some code and normally provided inside same plugin module.

\subsubsection{Payload}

@@

\subsubsection{Attributes}

@@

By convention keys are composed of two parts: name of MCC/Protocol at which Attribute was generated or must be consumed and name of Attribute itself separated by column. For example the Attribute with key HTTP:METHOD holds HTTP protocol method like GET, PUT, HEAD, etc. It is either generated by MCC implementing HTTP protocol or is filled by other code and is used by HTTP MCC to generate proper HTTP header.

Each MCC developed inside and outside the HED must have generated and consumed Attributes described in accompanying documentation.

\subsubsection{Security Attributes}

@@

@@

The Security Handlers are arranged into named queues. Elements in every queue are executed sequentially with Message as only argument. Different queues are executed at  different times. Which queue is executed at which case depends on MCC. Most MCCs implement two queues named "incoming" and "outgoing". The queue "incoming" is executed for Messages moving through the chain towards hosted application and passes as argument the Message with Payload of type corresponding to MCC type i.e. HTTP MCC passes Payload with parsed HTTP message. The "outgoing" queue is executed for Messages travelling to outside HED.

\subsection{Services and Clients}

The services are dynamically loaded on start up just like the MCCs. They almost identic to MCCs with an exception that they constitute the last link in the Message Chain. They are attached to the Chain in the same manner as other MCCs. But differently from MCC the \texttt{process} method of the Service does not pass Message to other components of the Chain. Instead they have to process incoming Messages and produce outgoing ones.

The clients are not represented by any specific component. The client code sees the Chain as single object with named entry points. Those entries are used by clients to insert request Message and get result Message on output. For simplifying task of writing clients there is library \texttt{arcclient} provided which wraps task of creating Chains and Messages for widely used SOAP, HTTP, TLS, TCP communications.

\subsection{Plexer}

In general case multiple Services living in the HED so the incomming Message should route to the proper Service. The Plexer MCC does this job. It takes the the ENDPOINT attributes of the message collected by other MCCs compares this attribute to regular expression defined in the configuration file and forward the message to the all matching service. It acts as a dispacher. The Plexer is also special in a sence that it is not a plugin but part of the \texttt{arcloader} library.

This Plexer provides only basic functionality and is capable of doing only simple routing. But because each MCC has multiple routing capabilities it is possible to provide pluggable MCC implementing more sophisticated and/or more specific routing algorithms.

\subsection{Error handling}

\texttt{Relevant classes: Arc::MCC\_Status}

For reporting errors each \texttt{process} method returns instance of MCC\_Status class. That object carries predefined set of common error codes.

This way for error reporting is mostly meant to be used for reporting problems related to code execution. For errors caused by processing corresponding protocol MCC should generate proper response Message which carries error description. Only if protocol does not provide error handling MCC\_Status should be used. It is also advisable to convert MCC\_Status error obtained from next MCC in the Chain into protocol psecific error Message if possible.

\subsection{Instantiation of the Chain}

\texttt{Relevant classes: Arc::Loader}

Chain instantiation is handled by Loader component. It takes XML configuration on input and then handles tasks of loading plugin libraries, identifying plugins in them, creating and linking objects of corresponding classes.

Each Loader object may create multiple non-intersecting chains and there may be multiple Loader objects in same executable. For each component mentioned in configuration Loader creates single object of corresponding class. On Loader destruction all handled components are destroyed too.

\subsection{Sessions and Contexts}

The HED defines few lifetimes for operations happening inside the Chain and components which can be associated with them:

%The HED provides few context containers associated with the Chain and the Message. Those are:

\item The Message lifetime - lasts as long as incoming and outgoing Messages are passing through the Chain forth and back. The Message it self in this case is used as container for associated components.

\item The Session lifetime - is defined by existence of some logical connection between Messages being processed. Corresponding container MessageContext is normally created by first MCC in a chain. Actual lifetime of that container is MCC specific. For TCP MCC it corresponds to TCP connection. The MessageContext holds objects inherited from MessageContextElement class. When Session ends all associated objects are destroyed.

%\item The Session Context. It holds objects inherited from MessageContextElement class and is persistent withion

\subsection{Alternative implementation languages}

In order to facilitate the developement of services, API bindings for languages other than C++ are provided and some service developement has already been done 
%Not much aboút the actual services, more about what they need to implement and what

The chain formulation using MCCs is up to the administrator.  Based on the configuration information the Loader components formulate the actual chain in the memory of the daemon.

\chapter{Implemented elements}

\section{Implemented MCCs}

\subsection{TCP MCC}

@@

\subsubsection{TLS MCC}
The server TLS MCC enables ...

@@

\subsection{HTTP MCC}

\subsection{SOAP MCC}

\section{Implemented Security Handlers}

\section{Implemented PDCs}

\chapter{Conclusion}
